---
title: "Practicum 1"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Question 1:

```{r}
life_exp = read.csv("/Users/sanikakalvikatte/Downloads/Life Expectancy Data.csv", header = T)
life_exp
```
Dropping Country, year and status as recommended in the Practicum.

```{r}
life_exp <- life_exp[ -c(1:3) ]
life_exp
```

# Question 2:

Exploring the data:

```{r}
summary(life_exp)
dim(life_exp)
```

Histogram of Life Expentancy:

```{r}
life_exp[!is.na(life_exp$Life.expectancy),]

```

# Question 3:

```{r}
library(dplyr)
library(ggplot2)
ggplot(life_exp, aes(Life.expectancy)) + 
  geom_histogram(aes(y = ..density..),color = "salmon", fill = "salmon") +
  geom_density(color = "blue", alpha = 10)
```

# Question 4:

```{r}
shapiro.test(life_exp$Life.expectancy)
ks.test(life_exp$Life.expectancy, "pnorm")
```

**Comment:** From the output, the *p-value < 0.05* implying that the distribution of the data are significantly different from normal distribution. In other words, we *cannot* assume the normality.

# Question 5:

```{r}
life_exp
```

```{r}
a <- life_exp$Life.expectancy
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
(a[o])
```
There are two outliers for life expectancy field.

```{r}
a <- life_exp$Adult.Mortality
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 40 outliers for adult morality field.

```{r}
a <- life_exp$infant.deaths
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 37 outliers for infant deaths field.

```{r}
a <- life_exp$Alcohol
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 3 outliers for alcohol field.

```{r}
a <- life_exp$percentage.expenditure
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 84 outliers for percentage expenditure field.

```{r}
a <- life_exp$Hepatitis.B
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 18 outliers for hepatitis field.

```{r}
a <- life_exp$Measles
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 48 outliers for measles field.

```{r}
a <- life_exp$BMI
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are no outliers for bmi field.

```{r}
a <- life_exp$under.five.deaths
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 84 outliers for under 5 deaths field.

```{r}
a <- life_exp$Polio
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 172 outliers for Polio field.

```{r}
a <- life_exp$Diphtheria
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 170 outliers for Diphtheria field.

```{r}
a <- life_exp$HIV.AIDS
m <- mean(a, na.rm = T)
sd <- sd(a, na.rm = T)
z <- (m - a)/sd
z <- abs(z)
o <- which(z>3)
length(a[o])
```
There are 69 outliers for HIV AIDS field.

# Question 6:

```{r}
newdf <- scale(life_exp)

life2 <- subset (life_exp, select = -Life.expectancy)
normalize <- function(x) 
  { 
    zscore <- (x - mean(x, na.rm = T))/sd(x, na.rm = T)
  }
life2 <- as.data.frame(lapply(life2, normalize))
life2$Life.expectancy <- life_exp$Life.expectancy
life2
life2 <- abs(life2)
life2
```

# Question 7:

```{r}
life_exp
x <- life_exp[-1]
x
y <- life_exp[1]
y
y$binary_expectency <- with(y, ifelse(Life.expectancy >= 70, "good", "not_good"))
y <- y[-1]
y
binary_expectency <- y
life2 <- cbind(life2, binary_expectency)
life2 <- subset(life2, select = -Life.expectancy)
life2
```

# Question 8:

```{r}
library(dplyr)
set.seed(300)
stratified <- y %>%
  group_by(binary_expectency) %>%
  sample_n(size=15, replace = T)

stratified

```

```{r}
library(caret)
set.seed(300)
#Spliting data as training and test set. Using createDataPartition() function from caret
indxTrain <- createDataPartition(y = y$binary_expectency,p = 0.15, list = FALSE)
training <- as.data.frame(y[-indxTrain,])
validation <- as.data.frame(y[indxTrain,])
head(training)
head(validation)
```

```{r}
# Rename columns
colnames(validation)[1] <- "validation"
validation
colnames(training)[1] <- "training"
training
```
```{r}
library(tidyr)
training %>% drop_na(training)
validation %>% drop_na(validation)
```

Performing data splitting with the sample method:

```{r}
#install.packages("sampler")
library(sampler)
library(splitstackshape)
library(zoo)
library(xts)
```


```{r}
mix <- runif(nrow(life2))

life2 <- life2[order(mix),]
life2

set.seed(300)
#created stratified sample
sampling <- stratified(life2,'binary_expectency', 0.15, bothSets = TRUE)
training <- as.data.frame(sampling[2])
validation <- as.data.frame(sampling[1])
```

```{r}
colnames(validation) <- gsub("SAMP1.", "", colnames(validation))
colnames(training) <- gsub("SAMP2.", "", colnames(training))
```

```{r}
training
validation
```


# Question 9:

```{r}
library("dplyr")
library("tidyr")

x["Adult.Mortality"][is.na(x["Adult.Mortality"])] <- 275
x["infant.deaths"][is.na(x["infant.deaths"])] <- 1
x["Alcohol"][is.na(x["Alcohol"])] <- 0.01
x["percentage.expenditure"][is.na(x["percentage.expenditure"])] <- 10
x["Hepatitis.B"][is.na(x["Hepatitis.B"])] <- 40
x["Measles"][is.na(x["Measles"])] <- 400
x["BMI"][is.na(x["BMI"])] <- 17
x["GDP"][is.na(x["GDP"])] <- 620
x["under.five.deaths"][is.na(x["under.five.deaths"])] <- 106
x["Polio"][is.na(x["Polio"])] <- 10
x["Diphtheria"][is.na(x["Diphtheria"])] <- 66
```

Load the training and test data

```{r}
#Loading the train data (x_label) and target data (y_label).

x_label <- c("Adult.Mortality",
"infant.deaths",
"Alcohol",
"percentage.expenditure",
"Hepatitis.B",
"Measles",
"BMI",
"under.five.deaths",
"Polio",
"Total.expenditure",
"Diphtheria")
y_label <- c("binary_expectency")
# x_label
# y_label <- as.data.frame(y_label)
# y_label
```

Select the value of K and calculate the ED

```{r}
#Calculating the euclidean distance

ED<-function(p,q){
distance=0
  for (i in (1:(length(p)-1))){
    distance=distance+(p[i]-q[i])^2
  }
  return(sqrt(distance))
}
```


```{r}
train_scaled <- as.data.frame(scale(training[,1:11], center = TRUE, scale = T))
validation_scaled <- as.data.frame(scale(validation[,1:11], center = TRUE, scale = T))
train_scaled$binary_expectency <- training[,y_label]
validation_scaled$binary_expectency <- validation[,y_label]
```


```{r}
knn_predict = function(y, train, k, x_label, y_label){
   # creating a new df - temp
  temp <- train
  #creating a new column - dist
  temp$dist <- apply(train[,x_label], 1, ED, y)
   #ordering the data
  train_ordered <- temp[order(temp$dist),]
  #filtering the data from 1 to value of k
  train_filterd <- train_ordered[1:k,]
  #data is filtered
  predicted_values <- tail(names(sort(table(train_filterd[, y_label]))), 1)
  return(predicted_values)
}
```

```{r}
#standardisation with z-score
z_scaler = function(df, col_name, x){
  mean_x <- mean(df[,col_name])
  sd_x <- sd(df[,col_name])
  return((x-mean_x)/sd_x)
}
y <- c(275,1,0.01,10,40,400,17,620,106,10,66)

#Adult Mortality = 275 | infant deaths = 1 | Alcohol = 0.01 | percentage expenditure = 10| Hepatitis B = 40 |Measles  = 400 | BMI = 17 | GDP = 620 | under-five deaths  = 106 | Polio = 10 | Diphtheria = 66
```

```{r}
#Standardization on y 
y_scaled <- c()
for (i in 1:length(y)){
  #print(training)
  #print(x_label[i])
  #print(y[i])
  y_scaled <- append(y_scaled, z_scaler(training, x_label[i], y[i]))
  #print(y_scaled)
}
#knn prediction
knn_predict(y_scaled, training, 5, x_label, y_label)
```


```{r}
train_scaled[,x_label]$Adult.Mortality[is.na(train_scaled[,x_label]$Adult.Mortality)] <- 275
train_scaled[,x_label]$infant.deaths[is.na(train_scaled[,x_label]$infant.deaths)] <- 1
train_scaled[,x_label]$Alcohol[is.na(train_scaled[,x_label]$Alcohol)] <- 0.01
train_scaled[,x_label]$percentage.expenditure[is.na(train_scaled[,x_label]$percentage.expenditure)] <- 10
train_scaled[,x_label]$Hepatitis.B[is.na(train_scaled[,x_label]$Hepatitis.B)] <- 40
train_scaled[,x_label]$Measles[is.na(train_scaled[,x_label]$Measles)] <- 400
train_scaled[,x_label]$BMI[is.na(train_scaled[,x_label]$BMI)] <- 17
train_scaled[,x_label]$under.five.deaths[is.na(train_scaled[,x_label]$under.five.deaths)] <- 620
train_scaled[,x_label]$Polio[is.na(train_scaled[,x_label]$Polio)] <- 106
train_scaled[,x_label]$Total.expenditure[is.na(train_scaled[,x_label]$Total.expenditure)] <- 10
train_scaled[,x_label]$Diphtheria[is.na(train_scaled[,x_label]$Diphtheria)] <- 66

train_scaled[,y_label][is.na(train_scaled[,y_label])] <- "good"

validation["Adult.Mortality"][is.na(validation["Adult.Mortality"])] <- 275
validation["infant.deaths"][is.na(validation["infant.deaths"])] <- 1
validation["Alcohol"][is.na(validation["Alcohol"])] <- 0.01
validation["percentage.expenditure"][is.na(validation["percentage.expenditure"])] <- 10
validation["Hepatitis.B"][is.na(validation["Hepatitis.B"])] <- 40
validation["Measles"][is.na(validation["Measles"])] <- 400
validation["BMI"][is.na(validation["BMI"])] <- 17
validation["GDP"][is.na(validation["GDP"])] <- 620
validation["under.five.deaths"][is.na(validation["under.five.deaths"])] <- 106
validation["Polio"][is.na(validation["Polio"])] <- 10
validation["Diphtheria"][is.na(validation["Diphtheria"])] <- 66

validation
validation <- validation %>% select(c(1:9,11))
validation
```
```{r}
y_scaled <- c(275,1,0.01,10,40,400,17,620,106,10,66)
```


```{r}

library(class)
knn(train = train_scaled[,x_label], test = y_scaled, cl = train_scaled[,y_label],k=5)
```

We can derive that, both the models give us same answer that is, the life expectency is good.

# Question 11:

```{r}
limit <- 1000
accuracy_scores <- c()
for (k in 2:10){
  predicted_value <- knn(train_scaled[1:limit,x_label], y_scaled, cl=train_scaled[1:limit,y_label],k=k)
  accuracy_score <- 100 * sum(y_scaled == predicted_value)/NROW(validation_scaled)
  accuracy_scores <- append(accuracy_scores, accuracy_score)
}
accuracy <- data.frame(accuracy_scores)
accuracy$k <- c(2:10)
accuracy
```

```{r}
ggplot(data = accuracy, aes(x = k, y = accuracy_scores)) +
geom_line(color = "blue") +
geom_point(color = "black") +
labs(title = "Plot of k from 2 to 10 versus accuracy" , x = "k",
y = "accuracy") + theme(plot.title = element_text(hjust = 0.5))
```


# Problem Set 2:

# Question 2:

```{r}
abalone = read.csv("/Users/sanikakalvikatte/Downloads/abalone.csv", header = T)
abalone
```

```{r}
target_data <- as.data.frame(abalone$Rings)
target_data
train_data <- abalone[,c(1:8)]
train_data
```

```{r}
df <- dummyVars(" ~ .", data=abalone)

encoded_data <- data.frame(predict(df, newdata=abalone))
encoded_data
```

```{r}
df <- dummyVars(" ~ .", data=train_data)

encoded_train <- data.frame(predict(df, newdata=train_data))
encoded_train

normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}
train <- as.data.frame(lapply(encoded_train, normalize))
train
```

# Question 4:

```{r}
# Euclidean distance
dist <- function(p,q)
{
d <- 0
for (i in 1:length(p)) {
d <- d + (p[i] - q[i])^2
}
dist <- sqrt(d)
}
```

```{r}
# Neighbors
neighbors <- function (train, new_data)
{
ds <- numeric(nrow(train_data))

# Calculate distance
for (i in 1:nrow(train_data))
{
p <- train[i,]
ds[i] <- dist(p,new_data)
}
ds <- as.vector(unlist(ds))
return(ds)
}
```

```{r}
# K-closest neighbors
k.closest <- function(neighbor,k)
{
ordered.neighbors <- order(neighbor)
kvalue <- ordered.neighbors[1:k]
}
```

```{r}
weight_prediction <- function(y, kclosest) {
length <- length(kclosest)
values <- y[kclosest,]
sums <- values[1]*2 + values[2]*1.5 + sum(values[3:length]*1)
w <- 2 + 1.5 + length(3:length)*1
wt_avg <- sums/w
}
```

```{r}
# KNN reg function
knn.reg <- function(new_data, target_data, train_data, k)
{
nbrs <- neighbors(train_data, new_data)
c <- k.closest(nbrs,3)
wt_mean <- weight_prediction(target_data, c)
return(wt_mean)
}
```


# Question 5:

```{r}
new_data <- c(0,0,1,0.34,0.491,0.245,0.4853,0.2532,0.0887,0.19)
#forecasting the price
knn.reg(new_data, target_data, train, 3)
```

# Question 6:

```{r}
df <- train
final_df <- target_data
head(df)
head(final_df)
```

```{r}
set.seed(300)

sample <- floor(0.15 * nrow(df))

train_set <- sample(seq_len(nrow(df)), size = sample)

data_train <- df[train_set,]

data_test <- df[-train_set,]

sample
```


```{r}
library(class)
train_labels <- final_df[train_set,1]
test_labels <- final_df[-train_set,1]
```

```{r}
knn_predict <- function(data_train, data_test, k){
knn_v <- knn(train = data_train,test = data_test,cl=train_labels, k)
mse_data <- data.frame(pred = as.numeric(levels(knn_v))[knn_v], actual = test_labels)
mse_data$e <- 0
for (i in 1:nrow(mse_data)) 
  {
  mse_data$e[i] <- mse_data$actual[i] - mse_data$pred[i]
  MSE <- mean(mse_data$e^2)
  }
return(MSE)
}
```

```{r}
k_values = c(1,2,3,4,5)
k_df <- data.frame(k_values)
k_df$mse_values <- 0

for (i in 1:nrow(k_df)){ 
k <- k_df$k_values[i]
k_df$mse_values[i] <- knn_predict(data_train,data_test,k)
}

k_df
```

# Problem 3:

# Question 1:

```{r}
house_prop <- read.csv("/Users/sanikakalvikatte/Downloads/raw_sales.csv")
head(house_prop)
```

```{r}
library("lubridate")
house_prop$datesold <- as_date(house_prop$datesold, "%YYYY%MM%DDT000000")
house_prop <- subset(house_prop,bedrooms != 0)
unique(house_prop$bedrooms)
house_prop
```
We need lubridate package to deal with dates here. 

```{r}
house_df <- house_prop %>% 
  mutate(year = format(house_prop$datesold, "%Y"), month = format(house_prop$datesold, "%m")) %>% 
  group_by (year, month) %>% 
  summarise(avg_price_br = mean(price/bedrooms))
```


```{r}
house_df
```

```{r}
house_df <- house_df[order(house_df$year, house_df$month),]
house_df$tper <- seq(1:nrow(house_df))
house_df
```

```{r}
#install.packages("xts")
library("xts")
house_df$date <- as.yearmon(paste(house_df$year, house_df$month, sep ="-"))
house_df
house_df <- as.data.frame(house_df)
```

# Question 3:

```{r}
mm <- as.list(house_df$month)

plot <- ggplot(house_df, aes(x = date, y = avg_price_br, length(40)))+
  geom_line(color ="salmon") + geom_point()
plot
```
Average price is increasing with respect date.

# Question 4:

```{r}
nrow(house_df)
o <- tail(house_df,n=4)
o
n = 150
last_4 <- house_df[n:(n-3),3]
last_4
w <- c(4,3,2,1)
sw <- w*last_4
weighted_moving_average <- sum(sw)/sum(w)
weighted_moving_average
```
The moving average weights for 4,3,2,1 are 207192.5 212319.9 215005.0 213527.0